### 2023-01-12 title-page.Rmd

---
 title: "Subjective status, Social Class and meritocracy perception"
 # author: Julio Iturra ^[Bremen International Graduate School of Social Sciences (jiiturra@uni-bremen.de)]
 bibliography: "input/bib/ess-merit.bib"
 csl: "input/bib/apa6.csl"
 css: "input/css/custom.css"
 output:
     bookdown::pdf_document2:
         template: default
         toc: false
         number_sections: false
     bookdown::html_document2:
          number_sections: false
     bookdown::word_document2:
          toc: true
 geometry: margin=1in
 header-includes:
   # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel} # para decimales con commas en vez de punto
   - \usepackage{times}           # Times New Roman
   - \usepackage{caption}
   - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
   - \usepackage{graphicx}
   - \usepackage{float}
   - \usepackage{booktabs}
   - \usepackage{longtable}
   - \usepackage{array}
   - \usepackage{multirow}
   - \usepackage{wrapfig}
   - \usepackage{float}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{tabu}
   - \usepackage{threeparttable}
 linestretch: '1.5'
 fontsize: 12pt
 link-citations: yes
 linkcolor: blue
 editor_options:
   chunk_output_type: inline
 # abstract: "Perceived meritocracy is understood as how societies describe the functioning of meritocracy, the ideal that promotes the distribution of resources in society based on individual effort and ability. Said this, are subjective and objective positions associated with perceived meritocracy? To what extent is objective position's influence on meritocracy perception moderated by subjective status? To answer these questions, we use nationally representative data from the 2018 wave of the Longitudinal Social Survey of Chile (N = 2,652). Regression analysis show that subjective social status has a robust positive association with perceived meritocracy, in line with the hypothesis. However, both positive and negative associations of income and occupational social class on perceived meritocracy were not confirmed. However, the hypothesis for education was confirmed, where higher educational attainment is associated with lower perceived meritocracy. Finally, we found that subjective status increases perceived meritocracy for those with higher incomes, tertiary education, and higher occupational social class."  
---

Julio Iturra^[Corresponding author], Bremen International Graduate School of Social Sciences (jiiturra@uni-bremen.de)

Juan Carlos Castillo, Department of Sociology, Universidad de Chile (juancastillov@uchile.cl)