### 2023-11-29 paper-blind.Rmd
ajustes apa y models

---
title: "Subjective status, Social Class and meritocracy perception"
linestretch: '1.5'          # interlineado
link-citations: yes         # citas linkeadas
fontsize: 12pt
linkcolor: blue
author:
- name: Julio Iturra ^[Corresponding author - jiturra@bigsss.uni-bremen.de]
  affiliation: Bremen International Graduate School of Social Sciences
- name: Juan Carlos Castillo
  affiliation: Universidad de Chile
bibliography: "input/bib/ess-merit.bib"
csl: "input/bib/apa6.csl"
css: "input/css/custom.css"
output:
  bookdown::html_document2:
    number_sections: false
  bookdown::word_document2:
    reference_docx: input/template.docx
    toc: false
    number_sections: false
  bookdown::pdf_document2:
    template: null
    toc: false
    keep_tex: false
    number_sections: false
    pandoc_args:
    - --template=input/mytemplate.tex #custom template para usar autores con afiliacion
geometry: margin=1in
header-includes:
  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel} # para decimales con commas en vez de punto
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \hypersetup {colorlinks = true, linkcolor = blue, urlcolor = blue}  
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
lang: en-US
abstract: "Perceived meritocracy is understood as how societies describe the functioning of meritocracy, the ideal that promotes the distribution of resources in society based on individual effort and ability. Said this, are subjective and objective positions associated with perceived meritocracy? To what extent is objective position's influence on meritocracy perception moderated by subjective status? To answer these questions, we use nationally representative data from the 2018 wave of the Longitudinal Social Survey of Chile (N = 2,652). Regression analysis show that subjective social status has a robust positive association with perceived meritocracy, in line with the hypothesis. However, both positive and negative associations of income and occupational social class on perceived meritocracy were not confirmed. However, the hypothesis for education was confirmed, where higher educational attainment is associated with lower perceived meritocracy. Finally, we found that subjective status increases perceived meritocracy for those with higher incomes, tertiary education, and higher occupational social class."  
keywords: subjective social status, meritocracy perception, social class
classoption: a4paper
editor_options: 
  chunk_output_type: console
---

```{r render,include=FALSE, eval=FALSE}
 rmarkdown::render("paper-blind.Rmd", output_format = "bookdown::pdf_document2")
 rmarkdown::render("paper-blind.Rmd", output_format = "bookdown::html_document2")
 rmarkdown::render("paper-blind.Rmd", output_format = "bookdown::word_document2")   
```

```{r setup, include=FALSE}
 knitr::opts_chunk$set(
 	echo = FALSE,
 	message = FALSE,
 	warning = FALSE,
 	cache = TRUE,
 	fig.align = 'center'
 )
 options(scipen=999)
 rm(list=ls())
 options(knitr.kable.NA = '')
```

* Word count: 6908

\pagebreak

# Introduction

Meritocracy is an ideal that promotes the distribution of resources in a society based on individual effort and ability [@Young1961]. In modern societies, merit is commonly embraced as a fair way of achieving status, in opposition to drawing on connections or inheritance [@Hadjar2008]. The increasing levels of economic inequality [@piketty_capital_2014] and the stagnation of social mobility [@oecd_broken_2018] has increased the attention of social scientist on the subjective aspects of meritocracy, mainly how societies perceive the functioning of meritocracy and its role in justifying economic inequality [@mijs_meritocracy_2020]. The literature on subjective meritocracy has pointed out the distinction between the descriptive dimension represented by perception (how it is) and preference as the normative dimension (how it should be) [@Castilloetal2019]. The role of social position has been claimed as a key factor in perceived meritocracy by social-psychological studies focusing on subjective social status [@bernardo_meritocracy_2021;@madeira_primes_2019]. In the sociological literature, attention has been paid to income and education as objective status characteristics [@Duru-bellat2012;@Castilloetal2019], but also little consideration of authority and control over labour represented by occupational social class [@Smith2012]. Both disciplinary approaches have been developing parallel research agendas on perceived meritocracy without an extensive dialogue.

Studies have shown the positive influence of economic self-interest [@morris_paradox_2022] but also the negative association between education [@Duru-bellat2012] and occupational social class [@Canales2015lse] with meritocracy perception. In other words, income increases the perception of meritocracy, but education and social class decrease it. Regardless of the materialist explanations of perceived meritocracy [@Goldthorpe2003], subjective social status and its anchoring in social comparison processes is another less-discussed social-psychological factor. The perceived standing in the social hierarchy of an individual or group depends on reference groups, such as top-bottom groups in society [@Evans2004] or their neighbours and local community [@condon_economic_2020], and it has been demonstrated that it can positively influence perceived meritocracy [@Castilloetal2019]. Furthermore, the reference group and reality blend (R&R) hypothesis [@Evans2004] has suggested that status perception and its discrepancies with objective position can be explained by the characteristics of reference groups involved in the comparison, arguing that status perception can lead to different outcomes than objective measures. Thus, what is the role of subjective social status on meritocracy perception? Furthermore, to what extent can the influence of objective position on perceived meritocracy be moderated by subjective status? We argue that subjective social status represents an independent process able to explain why some people perceive lower (or higher) meritocracy even when they hold (dis)advantageous positions in terms of income, educational attainment or occupational social class.

The present research aims to scrutinise the role of both subjective and objective social positions on perceived meritocracy. Regarding subjective social status, we claim that social comparison processes and reference groups play a crucial role in how individuals perceive meritocracy as an independent force. On the side of material factors, we examine the role of income and education as different status characteristics, but we also discussed the less addressed role of occupational social class. Finally, as an attempt to test the reference group and reality blend hypothesis, we tested the hypothesised positive moderating effect of subjective social status on the influence of objective status and social class.

## Perception of meritocracy

Merit is understood as the combination of individual effort and ability [@Sen2000]. Moreover, the meritocratic ideal assumes that status achievement is not constrained by how opportunities are structured in society [@mccall_exposure_2017]. Both constitute the substantive core of how meritocracy can be comprehended. However, studies on meritocracy have made undifferentiated use of concepts such as *Meritocratic Beliefs* [@mijs_how_2021a; @bernardo_meritocracy_2021] or *Perceived Meritocracy* [@Duru-bellat2012]. This lack of conceptual and empirical accuracy is often noticed in studies addressing meritocracy's subjective aspect without distinguishing between descriptive and normative dimensions [@Castilloetal2019]. The research on subjective meritocracy has focused on structural inequality as an explanatory factor, mainly analysing the role of social status, reporting mixed results given by different measurement strategies of the concept.

The concept of  _perception_ is understood as the subjective description of an object [@Janmaat2013]. In this regard, the descriptive dimension (*how is*) does not consider the normative evaluation of the object (*should be*) understood as preferences [@Janmaat2013]. In this regard, @Garcia2001 suggests perceived meritocracy is the degree to which individuals consider their society complies with the meritocratic principles. In other words, to what extent effort and talent are rewarded, indicating "how meritocracy is" rather than "how it should be". Previous evidence suggests that preferences, compared to perceptions, are more stable across status groups [@Castilloetal2019,p.10]. Thus, we expect that subjective and objective social positions can explain the differences in meritocracy perception.

## Subjective social status

The literature on subjective meritocracy has paid more attention to the structural position of individuals for explaning the emergence of meritocracy perceptions, assuming that structural positions and groups stratify [@castillo_perception_2022]. For example, perceived economic inequality in everyday life diminishes justification of inequality and increases demand for redistribution [@garcia-castro_perceiving_2020]. Nevertheless, it has been demonstrated that meritocracy primes reinforce negative evaluations of economically disadvantaged groups and strengthens internal attribution [@madeira_primes_2019].

Normative appraisals of qualitative or ascribed status characteristics such as gender or quantitative like income or educational level serve to explain status [@Jasso2001]. In this regard, social interactions serve as anchors through which individuals rationalise their societal position. The literature conceptualised it as subjective social status, understood as the perception of an individual or group's standing in the social hierarchy compared to a reference group, such as society itself [@slomczynski1987; @Evans2004] or local communities [@Condon2020]. A widely used measure is the MacArthur Scale of Subjective Social Status, in which individuals report social status on a ladder representing the strata of the social hierarchy [@adler_relationship_2000].

The literature has shown a persistent bias toward the middle of the hierarchy, where low-status individuals tend to overestimate their position while high-status individuals underestimate it [@Lindemann2014]. In this regard, previous theoretical approaches have suggested that individuals are likely to compare themselves with people or groups perceived as similar. These reference groups contribute to diminishing cognitive dissonance and seek to assemble better estimations about their position in society [@festinger1954theory; @Merton1968].

Social perception of meritocracy may be influenced by an individual's reference group and their perceived social standing within that group. The reference group and reality blend (R&R) hypothesis suggests that an availability heuristic may shape an individual's cognitive biases, reducing cognitive dissonance in situations with limited information [@KelleyEvans1995; @Evans2004]. Close reference groups can shape an individual's subjective estimates of their social standing, and the homophily within their social networks may influence their imagined ideas about society. For example, research has shown that individuals with networks of people with high-status occupations have a higher subjective social status in Chinese society, where social capital is important [@kim_social_2021]. Membership in organisations or participation in hobbies may also shape cognitive patterns of homophily and influence an individual's perceived social standing [@son_cognitive_2021].

The link of subjective status with attitudes toward inequality has been little studied. For example, @Vargas-Salfate2018 found that higher subjective status is associated with higher system justification. Similarly, @Schneider2015 evidence a positive association of subjective status with the perceived justice income inequality and higher internal poverty attributions. In addition, @Castilloetal2019 found that higher subjective status is related to higher perceived meritocracy. Based on this evidence, our main prediction is:

> $H_{\text{1}}$: subjective social status is positively associated with perceived meritocracy

## Income

According to the rational self-interest approach [@Meltzer81], higher-income individuals perceive higher meritocracy because they have accumulated economic resources through effort and talent. High status is also associated with lower inequality perception [@Evans2017] and higher justification of income inequality [@Trump2017].

There is evidence of a positive relationship between income and perceptions of meritocracy regarding the importance of individual merit in achieving success [@Reynolds2014; @Xian2017] and the extent to which intelligence and skills are rewarded [@Castilloetal2019; @Duru-bellat2012]. However, there is mixed evidence on how lower-income individuals rationalise inequality and meritocracy. Some research supports the "activated conflict theory," which suggests that the poor in unequal localities hold lower beliefs in meritocracy due to social comparisons that increase awareness of social status [@Newman2015]. Other research, based on the "relative power theory," challenges this finding, arguing that the rich hold more power due to income and wealth concentration, allowing them to disseminate legitimising ideas about economic inequality, and the poor are less critical of dominant beliefs like meritocracy as a psychological mechanism for coping with deprivation [@Solt2016]. However, other research shows that low-income individuals in highly unequal contexts have firmer meritocracy beliefs, but the differences tend to fade as inequality declines [@morris_paradox_2022].

Based on the literature, we expect that individuals with higher economic resources will attribute more importance to merit and therefore perceive that rewards in society are distributed according to it. Therefore, we propose the following hypothesis:

> $H_{\text{2a}}$: Income is positively correlated with perceived meritocracy.

## Education

As income reflects an individual's economic resources, studies on social stratification have raised the importance of educational attainment on status achievement [@Goldthorpe2003]. In this line, research on meritocracy perception has discussed two theoretical approaches. On the one hand, the *reproductionist* approach [@Bourdieu2009; @bernstein_class_2003] supports the *socialisation* hypothesis, suggesting school socialisation operates as a normative reinforcement of meritocratic ideals, then educated individuals are more willing to support meritocracy as a legitimising mechanism [@Lampert2013]. On the other hand, the *instruction* hypothesis suggests perceived meritocracy tends to weaken in individuals with higher educational attainment because they would be more aware of the influence of structural factors on status achievement [@Duru-bellat2012] considering experiences of educational mobility [@Canales2015lse] and higher chances to engage with complex political ideas [@bubak_perceptions_2019].

On the one hand, evidence in favour of the socialisation hypothesis has pointed out that education is associated with higher importance of individual factors in getting ahead in life [@Xian2017; @Reynolds2014; @mijs_paradox_2021], but also as a determinant of wages [@Kunovich2007]. On the other hand, evidence in favour of the instruction hypothesis shows that more educated individuals disagree that people are rewarded for their effort and talent [@Duru-bellat2012; @Castilloetal2019]. Additionally, the instructional hypothesis can be complemented by the fact that greater awareness of inequality may experience feelings of injustice and relative deprivation since individuals with the same educational attainment have access to unequal rewards [@Adams1965; @Smith-eta2020_relative-deprivation]. In this sense, it has been shown that higher educational attainment is associated with higher perceived inequality [@Castillo2012a_justice]. Thus, this evidence leads us to suggest the following:

> $H_{\text{2b}}$: education is negatively associated with perceived meritocracy.

## Occupational Social Class

The relationship between structural position and meritocracy has mainly focused on income and educational attainment, setting occupational social class aside from the discussion. Occupational class not only represent the synthesis of educational achievement and income levels [@Ganzeboom1992], but it also represents a symbol of power and authority in contemporary societies [@Chan2007]. Social class schemes draw on occupational characteristics, autonomy and authority in the workspace as meaningful distinctions to categorise individuals [@Erikson1992], making them relatively homogeneous groups [@Connelly2016]. Social class also comprise individuals' experience in the labour market regarding job opportunities and wages [@Svallfors2006]. For example, lower classes are exposed to higher labour market risk, such as the likelihood of job loss and lower wages, leading individuals to experience deprivation due to their risk exposure [@rehm_risks_2009].

The literature has paid little attention to occupational social class on meritocracy perception. For example, @Smith2012 found that individuals belonging to the petty bourgeoisie perceive higher levels of meritocracy than wage earners. These findings are supported by @Sandoval2017, which found that skilled and service workers attribute more importance to effort in achieving success than unskilled workers. However, @Canales2015lse found that upwardly mobile individuals in the service class perceived lower levels of meritocracy.

Even when individuals have improved their social class position from lower to non-manual or service classes [@Torche2005a], the stagnation of wages and persistent economic inequality [@EspinozaNunez2014] can lead to negatively perceived meritocracy due to frustrated economic prospects [@barozet_clase_2011] or by experiences of discrimination in the labour market [@cea_discriminacion_2020]. Achieving higher occupational social class positions may not always result in satisfactory outcomes, which can further negative meritocracy perceptions when confronted with labour market constraints [@Canales2015lse]. In this regard, we argue that occupational social class position contributes with information about the labour market, such as job opportunities and wages, leading us to suggest that:

> $H_{\text{2c}}$: Working in non-manual and service occupations is negatively associated with perceived meritocracy.

## Interplay of subjective and objective reality

The R&R hypothesis suggests that both subjective and material experiences interact with each other [@Evans2017]. Studies have shown that status and class positively influence subjective status, but it is less understood how subjective status moderates the effect of objective experiences on perceived meritocracy.

Higher-income individuals tend to have higher inconsistencies in perceived status [@Castillo2013]. For example, income, education, and occupational prestige have a positive non-linear relationship with subjective social status, with higher-status individuals showing more variation in subjective status [@Andersson2018a]. In addition, income level and educational attainment inconsistencies have been linked to political preferences [@Kreckel1993, @Lenski1967], and discrepancies between class identity and social class may also influence political preferences [@Sosnaud2013], but socio-demographic controls do not consistently support these results.

Studies on subjective social status as a moderating factor have primarily focused on its relationship to other subjective variables rather than how it interacts with income or education. One study using data from the International Social Survey Programme (ISSP) found that higher perceptions of inequality lead to higher preferences for redistribution and lower perceptions of meritocracy and that the relationship between inequality and meritocracy perception becomes weaker for those who perceive themselves as having a higher societal position [@Fatke2018]. Another study in European countries showed that economic inequality does not influence preferences for redistribution when controlled by perceived inequality but is positively moderated by subjective status [@Bobzien2019]. A study on bachelor students with high subjective social status found that those who perceive themselves as having a higher status and greater meritocracy tend to justify higher levels of economic inequality [@bernardo_meritocracy_2021].

Experiments have shown that manipulating individuals' perceptions of their social status can alter their attitudes towards inequality. For example, in Argentina, people in the lowest and highest income deciles tend to have the most bias, with those in the lowest stratum overestimating their position and those in the highest underestimating it. Manipulating this bias can change attitudes towards redistribution [@Cruces2011]. In Sweden, older, educated people with higher cognitive skills tend to be more accurate about their position in the income distribution. It is believed that those with strong meritocratic beliefs will not be affected by informational interventions and will have fewer preferences for redistribution [@karadja_richer_2017]. Reference groups can also influence perceptions of distributive justice. Misperception is generally higher in large reference groups than in small ones. When informed of their actual position, those who overestimate their position around the median of any reference group, especially those in the lowest income stratum, tend to increase their perception of inequality and injustice [@hvidberg_social_2020].

Studies have demonstrated that three factors influence perceptions of social status: socioeconomic status, economic and social capital, and moral perceptions of individual achievement [@macclure_naming_2020]. Individuals who identify as lower status tend to have low incomes, less education, and live in disadvantaged areas. However, some individuals who belong to middle or high socioeconomic strata may still identify as lower status due to living in affluent neighbourhoods or having higher educational credentials, which symbolise higher social and cultural capital [@macclure_naming_2020]. Those who identify as high status tend to consistently have higher socioeconomic status, with educational meritocracy often serving as a criterion for their social position [@macclure_naming_2020].

Subjective social status can reflect an individual's position in the social structure and their experience of economic inequality through comparisons with nearby reference groups. Therefore, we predict that subjective status differences can lead to perceived meritocracy differences within socioeconomic groups. Thus, we predict that:

> $H_{\text{3}}$: The association of objective status and social class on perceived meritocracy is positively moderated by subjective social status.

To summarise our hypothesis, Figure \@ref(fig:tmodel) depicts the theoretical model proposed in the study.

```{r tmodel, fig.cap='Theoretical model',out.height='30%', fig.align='center'}
 knitr::include_graphics(path = "input/images/scheme.PNG")
```

\pagebreak
# Data, variables and methods

_Data_

The data used correspond to the third wave of the Longitudinal Social Study of Chile [@coes_estudio_2018], a survey developed to analyse the evolution of conflict and social cohesion in Chilean society over time. The sampling is probabilistic, stratified, clustered and multistage, with 3748 participants aged between 18 and 75. An analytical sample of 2652 cases with complete information is used for the analyses.

_Variables_

The study variables and their description can be seen in Table \@ref(tab:desc01). The main dependent variable is _perceived meritocracy_, operationalised through the average score of two indicators (_r_ = 0.76) that capture the degree of agreement on whether effort and intelligence are rewarded in Chile and that have been used in international comparative studies [@Duru-bellat2012]. Both indicators are five-category Likert scales, where higher values represent higher perceptions of meritocracy.

*Subjective social status* is measured using an 11-point scale variable in which respondents identify themselves on the social ladder by answering _"In our society, some groups are usually at the higher levels, and others tend to be placed at the lower levels of society. Using the presented scale, where 0 is the lowest level, and 10 is the highest level, where would you put yourself in Chilean society?."_, where higher values represent a higher perceived social status.

The objective position is measured through three variables. First, _Household Income Decile_ is calculated using monthly per capita household income and categorised into deciles. Second, _Educational level_ corresponds to the five categories for educational attainment following the International Standard Classification of Education [@unesco_international_2012].

Control variables are age and gender

_Method_

Considering the dependent variable as a continuous, stepwise ordinary least square regression models are used (OLS). First, models are estimated for each of the independent variables. Second, a series of estimated models consider the interaction between objective position and subjective status. Finally, interaction effects are presented using conditional linear predicted values <!-- and linear predictions -->.

# Results

## Descriptive

```{r plot-likert, fig.height=3, fig.width=10, fig.cap='Perceived meritocracy items'}
 load(here::here("input/data/proc/elsoc_w03.RData"))   
 pacman::p_load(dplyr,sjPlot,ggplot2)
 elsoc_w03$Q05h[elsoc_w03$Q05h=="No responde"] <- NA
  elsoc_w03$merit <-(as.numeric(elsoc_w03$meffort)+as.numeric(elsoc_w03$mtalent))/2
  elsoc_w03<- elsoc_w03 %>%
    dplyr::select(meffort,mtalent,merit,income=inc10h, edcine2, ess, sexo, edad,ponderador02) %>% 
        na.omit() 
 dat_likert_s1 <-
   dplyr::select(elsoc_w03,meffort,mtalent)
 dat_likert_s1 <- sjmisc::rec(dat_likert_s1, rec="rev", append=FALSE)

   ggplot2::theme_set(ggplot2::theme(panel.background = ggplot2::element_rect(fill = "gray85",
                                                                              colour = "gray85"),
                                     panel.border = ggplot2::element_blank(),
                                     axis.text.y = ggplot2::element_text(size = 13,
                                                                         hjust = 1),
                                     title = ggplot2::element_text(size = 13,
                                                                   face = "bold"),
                                     legend.text = ggplot2::element_text(size = 12),
                                     plot.caption = ggplot2::element_text(size = 10,
                                                                          face = "plain",
                                                                          hjust = 1)))

     sjPlot::plot_likert(dplyr::select(dat_likert_s1,meffort_r,mtalent_r),
                         geom.colors = "PuBu",
                         geom.size = 0.8,
                         axis.labels = c("People are rewarded for their efforts",
                                         "People are rewarded for their intelligence and ability"),
                         cat.neutral = 3,
                         grid.range  =  c (1.2 , 1.2),
                         values  =  "sum.outside",
                         reverse.scale = F,
                         show.n = FALSE) +
       theme(legend.position="bottom") +
       ggplot2::scale_fill_manual(
         values = c("#b3b3b3ff", "#f1eef6ff", "#bdc9e1ff","#74a9cfff","#0570b0ff"),
         breaks=c("neutral",4,3,2,1),
         labels=c("Neither agree \nnor disagree","Strongly \ndisagree", "Disagree",
                     "Agree", "Strongly \nagree")) +
       labs(caption = paste0("Source: Authors calculation based on ELSOC 2018",
                             " (N = ",dim(elsoc_w03)[1],")"))
```

Figure \@ref(fig:plot-likert) depicts the distribution of responses of the two indicators for the perception of meritocracy, showing that most individuals tend to disagree or strongly disagree with the statement. For example, regarding the statement about how _effort_ is rewarded, 50.8% answered that they disagree or strongly disagree. In the same direction, the perception regarding rewards from  _intelligence and ability_, the level of disagreement decreases to 42.6%. Finally, about 22% of the respondents do not have a negative or positive perception of the link between rewards and merit in society.

```{r corr-mat,  fig.cap='Correlation matrix between social status and perceived meritocracy', fig.height=8, fig.width=10, out.width='80%'}
 load(here::here("input/data/proc/elsoc_w03.RData"))
 pacman::p_load(dplyr,sjPlot,ggplot2,psych,polycor)
 elsoc_w03$merit <-(as.numeric(elsoc_w03$meffort)+as.numeric(elsoc_w03$mtalent))/2
  df_corr<- elsoc_w03 %>%
    dplyr::select(merit,income=inc10h, edcine2,ess, sexo, edad) %>% 
        na.omit() %>% 
    dplyr::select(-sexo,-edad) 
 cormat<- polycor::hetcor(data = df_corr,parallel = T)
 cormat <- as.matrix(cormat)

  diag(cormat) = NA #set diagonal values to NA
  # Set Row names of the matrix
  rownames(cormat) <- c("A. Meritocracy perception",
                        "B. Household income decile",
                        "C. Educational level",                        
                        "D. Subjective social status"
                        )
  #set Column names of the matrix
  colnames(cormat) <-c("(A)", "(B)","(C)","(D)")
  #Plot the matrix using corrplot
  corrplot::corrplot(cormat,
    method = "color",
    addCoef.col = "#000390",
    type = "upper",
    tl.col = "black",
    col=colorRampPalette(c("white","#0068DC"))(12),
    bg = "white",
    na.label = "-")
  text(1, 0.5, paste0("Source: Authors calculation based on ELSOC 2018",
                             " (N = ",dim(df_corr)[1],")"),   
       cex = 1)

```

Figure \@ref(fig:corr-mat) shows a positive correlation between perceived meritocracy and subjective status, with a slightly stronger association for the indicator for effort (r = 0.08) compared to intelligence (r = 0.06). Perceived meritocracy has a weak positive correlation with income decile, but weak negative correlations with education and social class.

## Multivariate analysis

```{r ols-merit, echo=FALSE, results='asis'}
pacman::p_load(texreg,dplyr) 
load(here::here("input/data/proc/elsoc_w03.RData"))
elsoc_w03$merit <- (as.numeric(elsoc_w03$meffort)+as.numeric(elsoc_w03$mtalent))/2 
dfreg <- elsoc_w03 %>%
  dplyr::select(merit,income=inc10h, edcine2, ess, sexo, edad,ponderador02) %>%
  na.omit()

dfreg$income[dfreg$income=="99"] <- NA
dfreg$ess_cwc <- mean(dfreg$ess) - dfreg$ess

m_inc <- lm(merit~income,data = dfreg,weights =ponderador02)
m_educ <- lm(merit~edcine2,data = dfreg,weights =ponderador02)
m_status <- lm(merit~income+edcine2+ess+sexo+edad,data = dfreg,weights =ponderador02) 
m_inter_inc <- lm(merit~income*ess+edcine2+ess+sexo+edad,data = dfreg,weights =ponderador02)
m_inter_edu <-lm(merit~income+edcine2*ess+ess+sexo+edad,data = dfreg,weights =ponderador02) 
models <- list(m_inc,m_educ,m_status,m_inter_inc,m_inter_edu)
caption <- "Linear Regression for Perception of Meritocracy"

knitreg(
  l = models,
  caption = paste("(\\#tab:merit)",caption),
  caption.above = T,
  custom.coef.map = list(
    income = "Income Decile",
    edcine2cine12 ="Primary & Lower secondary",
    edcine2cine34 = "Upper secondary",
    edcine2cine5 = "Short-cycle tertiary",
    edcine2cine678 = "Tertiary or higher",
    ess="Subjective Social Status",
    "income:ess" = "Income x SSS",
    "edcine2cine12:ess"= "Primary & Lower secondary x SSS",
    "edcine2cine34:ess"= "Upper secondary x SSS",
    "edcine2cine5:ess"= "Short-cycle tertiary x SSS",
    "edcine2cine678:ess"= "Tertiary or higher x SSS"
  ),
  groups = list("Education (ref: Incomplete Primary or less)"=2:5)
)
```

According to model 5, subjective social status has a strong, positive, and statistically significant association with perceived meritocracy in all specifications, supporting the reference group hypothesis (H1). Individuals with higher subjective social status tend to perceive higher meritocracy when comparing themselves with lower or higher groups in society and estimating their position on the social ladder.

Model 5 in Table \@ref(tab:merit) shows that household income has a positive but insignificant effect, therefore not supporting the rational self-interest hypothesis (H2a). Results show that as educational level increases, perception of meritocracy decreases. The evidence supports the instruction hypothesis (H2b) that people with higher education tend to view meritocracy more critically. Short-cycle tertiary education individuals perceive meritocracy slightly more than those with tertiary or higher education. Finally, Model 5 shows a negative effect of non-manual and service classes on social class, but the association is not statistically significant, therefore not supporting the labour market position hypothesis (H2c).

Table \@ref(tab:merit) shows the interaction between subjective social status, objective status, and social class in models 6 to 8, consistent with the hypothesized interaction of reference groups and objective reality (H3).

```{r int-income, fig.cap='Conditional effects of income decile by subjective social status on perceived meritocracy', fig.width=10, fig.height=5}
set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1")
options(ggplot2.discrete.fill = set1)
options(ggplot2.discrete.colour = set1)
ggplot2::theme_set(theme_minimal())
ggplot2::theme_update(text=element_text(size=15,  family="serif")) 
pacman::p_load(texreg,dplyr,marginaleffects,ggplot2) 
load(here::here("input/data/proc/elsoc_w03.RData"))
elsoc_w03$merit <- (as.numeric(elsoc_w03$meffort)+as.numeric(elsoc_w03$mtalent))/2

elsoc_w03$ess_cwc <- sjmisc::center(x = elsoc_w03$ess)
summary(elsoc_w03$ess_cwc)

dfreg <- elsoc_w03 %>%
  dplyr::select(merit,income=inc10h, edcine2, ess=ess_cwc, sexo, edad,ponderador02) %>%
  na.omit()
ess_max<- mean(dfreg$ess) + sd(dfreg$ess)
ess_min<- mean(dfreg$ess) - sd(dfreg$ess)
ess_mea<- mean(dfreg$ess)

dfreg$income[dfreg$income=="99"] <- NA
m_inter_inc <- lm(merit~income*ess+edcine2+ess+sexo+edad,data = dfreg,weights=ponderador02)

df_pref <-
  predictions(
    m_inter_inc,newdata = datagrid(
      ess = c(ess_min, ess_mea, ess_max),
      income = 1:10,
      edcine2 = "cine34",
      sexo = "Female",
      edad = mean(dfreg$edad)
    )
  )

df_pref$ess <- as.factor(df_pref$ess)
df_pref %>% 
ggplot(aes(y=estimate,x=income, fill=ess,color=ess,ymin=conf.low, ymax=conf.high)) +
  geom_line(aes(y=estimate,x=income),size=0.75) +
  geom_point(size=2) +
  geom_ribbon(alpha=0.2,size=0.5,linetype=1) +
  scale_color_discrete(name="Sub.Status",labels = c("-1SD","Mean","+1SD")) +
  scale_fill_discrete(name="Sub.Status",labels = c("-1SD","Mean","+1SD"))+
   labs(y = "Perceived Meritocracy",x="Income Decile",) +
   ylim(1, 5) +
   scale_x_continuous(breaks = 1:10,labels = paste0("D",1:10)) +
   theme(legend.position="right",
         legend.direction = "vertical",
         axis.text=element_text(size=15),
         legend.background = element_blank(),
         legend.box.background = element_rect(colour = "black"))
```

Model 6 shows that subjective social status positively moderates the association of household income with perceived meritocracy. Figure \@ref(fig:int-income) depict that this association is lower for those with low subjective social status and higher for those with high subjective social status. Subjective social status does not significantly differentiate perceived meritocracy for individuals with household income below Decile 5 but does for those with income above Decile 6. This result suggests that individuals with household income above the median and high subjective social status perceive that merit is rewarded in society.

Model 7 shows that subjective social status positively moderates the negative influence of educational level on perceived meritocracy, particularly among tertiary education groups. Figure \@ref(fig:int-educ) shows slight differences in perceived meritocracy among groups with secondary or less education based on subjective social status. However, for short-cycle and tertiary education groups, perceived meritocracy is higher for those with high subjective social status and lower for those with low subjective social status. The negative association of education can be observed among individuals with tertiary education who perceive themselves as close to the bottom of society and have a critical view of how merit is rewarded.

```{r int-educ, fig.cap='Conditional effects of Education by subjective social status on perceived meritocracy', fig.width=10, fig.height=5}
set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1")
options(ggplot2.discrete.fill = set1)
options(ggplot2.discrete.colour = set1)
ggplot2::theme_set(theme_minimal())
ggplot2::theme_update(text=element_text(size=15,  family="serif")) 
pacman::p_load(texreg,dplyr,marginaleffects,ggplot2,stringr) 
load(here::here("input/data/proc/elsoc_w03.RData"))
elsoc_w03$merit <- (as.numeric(elsoc_w03$meffort)+as.numeric(elsoc_w03$mtalent))/2 
elsoc_w03$ess_cwc <- sjmisc::center(x = elsoc_w03$ess)

dfreg <- elsoc_w03 %>%
  dplyr::select(merit,income=inc10h, edcine2, ess=ess_cwc, sexo, edad,ponderador02) %>%
  na.omit()
dfreg$ess_cwc <- mean(dfreg$ess) - dfreg$ess
ess_max<- mean(dfreg$ess) + sd(dfreg$ess)
ess_min<- mean(dfreg$ess) - sd(dfreg$ess)
ess_mea<- mean(dfreg$ess)
dfreg$income[dfreg$income=="99"] <- NA
m_inter_edu <- lm(merit~income+edcine2*ess+ess+sexo+edad,data = dfreg,weights=ponderador02)

df_pref <-
  predictions(
    m_inter_edu,newdata = datagrid(
      ess = c(ess_min, ess_mea, ess_max),
      income = 6,
      edcine2 = levels(dfreg$edcine2),
      sexo = "Female",
      edad = mean(dfreg$edad)
    )
  )

df_pref <- as.data.frame(df_pref)
df_pref$ess <- as.factor(df_pref$ess)

df_pref$edcine2 <- factor(df_pref$edcine2,levels = ,labels = c("Incomplete Primary or less","Primary & Lower secondary","Upper secondary ","Short-cycle tertiary","Tertiary or higher"))

df_pref %>% 
ggplot(aes(y=estimate,x=edcine2, color=ess, group=ess,ymin=conf.low, ymax=conf.high)) +
  geom_line(size=0.75) +
  geom_point(size=2) +
  geom_errorbar(alpha=1,size=1,width=.1,linetype=1) +
  scale_color_discrete(name="Sub.Status",labels = c("-1SD","Mean","+1SD")) +
  scale_fill_discrete(name="Sub.Status",labels = c("-1SD","Mean","+1SD"))+
  labs(y = "Perceived Meritocracy",x="Education") + 
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
     theme(legend.position="right",
         legend.direction = "vertical",
         axis.text=element_text(size=15),
         legend.background = element_blank(),
         legend.box.background = element_rect(colour = "black"))
```

<!-- Esto dice que en el fondo el efecto negativo del nivel educacional es mucho más marcado en aquellos que perciben un estatus más bajo.   -->

\pagebreak

# Dicussion

The reference group hypothesis (H1) proposes that an individual's perceived position on the social ladder influences their perception of meritocracy. Specifically, how individuals rank themselves in relation to the top and bottom groups in society can impact their perception of meritocracy. For example, high-status groups are typically associated with high-income, educated, or prestigious occupations in society [@condon_economic_2020] and may be seen as deserved winners. Our results show that when individuals perceive themselves as being closer to high-status groups, they tend to agree that effort and talent are rewarded. These findings are consistent with previous research on individualistic poverty attributions [@Schneider2015] and perceived meritocracy [@Castilloetal2019]. In general, the results align with the prediction of the self-interest hypothesis, but perceived status is more important than material reality in explaining perceived meritocracy.

The results did not support the self-interest hypothesis (H2a), which predicts that material factors such as household income would significantly influence individuals' perceptions of meritocracy. Specifically, household income was found to have no significant impact on perceived meritocracy. Previous research has suggested that the positive relationship between income and perceived meritocracy may be due to a within-group social comparison effect, as the association disappears when controlling for income distance from the average of an individual's educational group [@Castilloetal2019]. However, our findings contradict previous evidence that poor individuals are more likely to support meritocracy in highly unequal contexts [@morris_paradox_2022;@Solt2016], as we did not observe significant differences in perceived meritocracy between individuals with low and high income.

Our results support the instruction hypothesis (H2b) prediction, which posits that higher levels of educational attainment are associated with lower levels of perceived meritocracy. This is in contrast to the socialisation hypothesis, which suggests that individuals internalise norms about meritocracy through socialisation [@Lampert2013]. Furthermore, our findings suggest that higher levels of education may lead to more critical views of the distributive system and the idea that effort and talent are rewarded in society rather than reinforcing meritocratic preferences, as suggested by @Duru-bellat2012.

The results do not support the labour market hypothesis (H2c), which predicts that non-manual and service social classes will have lower levels of perceived meritocracy. This suggests that labour market position does not influence individuals' perceptions of meritocracy. While prior research has suggested that social class contributes to experiences in the labour market [@Svallfors2006] and that negative job experiences or wages may lead to lower perceptions of meritocracy [@Canales2015lse], the lack of significant differences among social classes in our study may be due to the lack of impact of income as a relevant labour market outcome.

<!-- *Reference groups and reality blend* -->
As an empirical attempt to scrutinise the Reference Group and Reality blend (R&R) hypothesis, we tested the moderating role of subjective status on objective reality. The results align with previous evidence on the moderating role of subjective social status on objective [@Bobzien2019] or perceived economic inequality [@Fatke2018]. Furthermore, we found that the conditional effect of status and class on perceived meritocracy can be observed among those on the top of the social hierarchy. In addition, apart from income levels, we took a step forward and explored the role of educational status and social class.

As suggested by the R&R approach, individuals who clearly understand their higher status/class positions tend to perceive higher levels of meritocracy following the principles of meritocracy. However, those with higher structural positions but lower subjective status tend to be more sceptical of meritocracy. Previous research [@kim_social_2021] has indicated that social comparison through limited information resulting from homophily in social networks may lead to biased perception, but this assumption cannot be confirmed with the current data. Our findings align with previous research on the impact of status perception can lead to attitudinal change [@hvidberg_social_2020;@Cruces2011]. Nevertheless, it is important to exercise caution when interpreting the correlational nature of our evidence.

One possible explanation for the observed patterns in our data is that social comparison processes [@festinger1954theory] may lead to subjective experiences of deprivation, even among those who are objectively at the top of the social hierarchy. As proposed by @Condon2020, upward comparisons may create anxiety among individuals of all status groups, resulting in deprivation, lower subjective status, and an increased preference for redistribution. Additionally, the Inequity theory [@Adams1965] posits that feelings of being under-rewarded can lead to a sense of deprivation. In the context of meritocracy, merit can be understood as a representation of time and energy investment. Thus, it is expected that those who perceive themselves as lower in the social hierarchy will also perceive lower levels of meritocracy, especially among those with affluent, educated, and highly skilled jobs. While these groups may be objectively successful, some individuals may still perceive themselves as lower in society and hold a critical view of meritocracy.

# Conclusion

One limitation of our study is using a two-item indicator to measure perceived meritocracy, which could be improved upon in future research. Additionally, operationalising the "bottom" and "top" groups in society to characterise reference groups could be more accurate if it considers other status characteristics such as gender, age, and migration status. It is also important to note that the correlational design of the study does not allow for causal conclusions. Future research using a panel design in the ELSOC survey will be able to better examine the relationship between structural change and its impact on perceived meritocracy.

Other areas for future research include the role of social network composition in understanding the influence of the closest reference group on status perception, the direct and moderating effects of subjective social status perceptions and preferences towards inequality, particularly to preferences for redistribution and justification of salary gaps, and the behaviour of individuals at the top of the social hierarchy who perceive themselves as lower in the ladder in regards to civic engagement, political participation, and altruistic behaviour.

# References

::: {#refs}
:::

<!-- Although <div> is an HTML tag, this method also works for other output formats such as PDF. -->

# (APPENDIX) Appendix {.unnumbered}

```{r,include=FALSE}
 # Funciones para renderizar tablas tanto en HTML como en pdf
 if (!require("pacman")) install.packages("pacman")  #si falta pacman, instalar
 if (!require("tinytex")) install.packages("tinytex")#si falta tinytex, instalar
 pacman::p_load(knitr, summarytools, kableExtra, dplyr, lavaan,
                haven, htmlwidgets, webshot, sjPlot,sjlabelled) # librerias
 knitr::opts_chunk$set(warning = FALSE,  # mensaje de warning
                       message = FALSE,  # mensajes/avisos de librerias
                       cache = FALSE,    # cache de los chunks,usar analisis pesados
                       out.width = '85%',# largo de imagen en %
                       fig.pos= "H",     # posicion figuras H = HERE
                       echo = FALSE      # incluir chunk en output
 )
 # Otras configuraciones
 options(scipen=999) # notacion cientifica
 options(knitr.kable.NA = '') # NA en kable = ''
 # Css
 st_css()
 # Cargar base de datos procesada
  load(here::here("input/data/proc/elsoc_w03.RData"))
  elsoc_w03$meffort <-
   factor(elsoc_w03$meffort,
          levels = 1:5,
          labels = c("Strongly disagree","Disagree",
                     "Neither agree nor disagree", "Agree", "Strongly agree"))

  elsoc_w03$mtalent <-
   factor(elsoc_w03$mtalent,
          levels = 1:5,
          labels = c("Strongly disagree","Disagree",
                     "Neither agree nor disagree","Agree","Strongly agree"))

  elsoc_w03$edcine2 <-
   factor(elsoc_w03$edcine2,levels = levels(elsoc_w03$edcine2),
          labels = c("Incomplete Primary or lower","Primary & Lower secondary",
            "Upper secondary ","Short-cycle tertiary","Tertiary or higher"))
  elsoc_w03$edcine2 <- sjlabelled::set_label(elsoc_w03$edcine2,"Education")

  elsoc_w03$inc10h[elsoc_w03$inc10h=="99"] <- NA
  elsoc_w03$inc10h <- sjlabelled::set_label(elsoc_w03$inc10h,"Income Decile")

  elsoc_w03$merit <- 
   (as.numeric(elsoc_w03$meffort)+as.numeric(elsoc_w03$mtalent))/2

  elsoc_w03$merit <- sjlabelled::set_label(elsoc_w03$merit,"Perceived meritocracy")
  df_desc<- elsoc_w03 %>%
    dplyr::select(merit,meffort,mtalent,
           ess,inc10h,edcine2,
           sexo,edad) %>% 
    dplyr::select(-meffort,-mtalent,
                  # -sexo,-edad,
                  # -ppolcat,-lngap_perc
                  )
  varlabs <- sjlabelled::get_label(df_desc) #recuperar labels
  df_desc <- na.omit(df_desc)
  set_label(df_desc) <- varlabs  
 st_options(lang = "en",
            footnote = paste0("Source: Authors calculation based on ELSOC 2018 (N = ",
            dim(df_desc)[[1]],")"),
            bootstrap.css = F,
            # custom.css = "input/css/dfsummary.css"
            )
 df<- dfSummary(df_desc,
                plain.ascii = FALSE,
                style = "grid",
                tmp.img.dir = "/tmp",
                graph.magnif = TRUE,
                headings = F,  # encabezado
                varnumbers = F, # num variable
                labels.col = T, # etiquetas
                na.col = F,    # missing
                graph.col = T, # plot
                valid.col = T,max.string.width = 50,
                max.distinct.values = 8 # n valido 
                # col.widths = c(1000,10,10,10)
                )
 df$Variable <- NULL # delete variable column
 df$Valid <- NULL
 summarytools::view(df,file = "output/tables/desc01.html") # Ver tabla en un archivo HTML
 webshot(url ="output/tables/desc01.html",
         file ="output/tables/desc01.png",
         vwidth = 700,vheight = 600) # Remover archivo HTML
```

```{r desc01, echo=FALSE, results='asis',fig.align='center'}
 # REVISAR EL N
 knitr::kable(data.frame(image='![](output/tables/desc01.png){height=80%}'),    # generar tabla
              caption="Study variables",
             col.names = "",
             row.names = F,
             format = "pandoc") %>%
  kable_styling(latex_options = c("HOLD_position","scale_down"),
                full_width = fwidth)
```